\documentclass{article}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{hyperref}

\title{User Manual}
\date{}

\begin{document}

\maketitle
\tableofcontents

\section{Overview}
This document explains how to use The Sleeping Lion's graphical interface to its best.

\section{Shortcuts}
There are several useful shortcuts implemented in The Sleeping Lion's graphical interface to allow faster editing.

\subsection{Enter}
Pressing \textit{Enter} will refresh the current card's preview.

\subsection{Ctrl+Mouse Wheel}
Using \textit{Ctrl+Mouse Wheel} allows to zoom in or out the card preview.

\subsection{Ctrl+Tab and Ctrl+Shift+Tab}
\textit{Ctrl + Tab} allows to switch the widget focus in the current tab. This shortcut allows to iterate through the entries in a tab, for example going from the \textit{Card name} entry to the \textit{Level} entry. When reaching the last entry in a tab, \textit{Ctrl + Tab} will loop back to the first entry.

\textit{Ctrl + Shift + Tab} iterates through the entries in reverse order, from bottom to top. When reaching the first entry, it loops back the the last one.

\subsection{Alt+Right Arrow and Alt+Left Arrow}
\textit{Alt+Right Arrow} allows to iterate through the different tabs in the graphical interface. This shorcut will focus on the tab immediately to the right of the current tab. When reaching the last one, this shorcut will loop back to the first one.

\textit{Alt+Left Arrow} does the same thing but in reverse order, from the current tab to the one immediately before it.


\end{document}

